---
title: "grmpyTemplate Documentation"
output: word_document
---

To create a representative midpoint template, we want to insure: 
1) An equal representation from each age range
2) An equal distribution of males and females

###Locate demographic data of interest:

PNC data from timepoint 1 (ses_01):/data/joy/BBL/studies/pnc/n1601_dataFreeze/demographics/n1601_demographics_go1_20161212.csv

grmpy data from timepoint 2 (ses_02):/data/jux/BBL/studies/grmpy/rawPsycha1/old/demographics_20180803.csv

```{r include=FALSE}
library(dplyr)
library(sqldf)

#Map to and load metadata

grmpyTimepoint1_dir <- "/data/joy/BBL/studies/pnc/n1601_dataFreeze/demographics/"
grmpyTimepoint2_dir <- "/data/jux/BBL/studies/grmpy/rawPsycha1/old"

grmpyTimepoint1_fp <- file.path(grmpyTimepoint1_dir, "n1601_demographics_go1_20161212.csv")
grmpyTimepoint2_fp <- file.path(grmpyTimepoint2_dir, "demographics_20180803.csv")

T1_metadata <- read.delim(file = grmpyTimepoint1_fp, header = T, as.is = T, sep = ",")
T2_metadata <- read.delim(file = grmpyTimepoint2_fp, header = T, as.is = T, sep = ",")

#Select variables of interest (in this case:L bblid, sex, age at time of scan)

T1 <-T1_metadata %>%
  select(bblid, sex, ageAtScan1)
T2 <- T2_metadata %>%
  select(bblid, sex, visitagemonths)

#Filtering PNC data for grmpy subjects

T1_grmpySubjectsExtracted <- sqldf("SELECT bblid, ageAtScan1 
              FROM T1
              JOIN T2 USING(bblid)")

grmpyTemplateDemographics <- merge(T2, T1_grmpySubjectsExtracted, by= "bblid")

```

For subjects that overlap between PNC and grmpy datasets (now collected in grmpyTemplateDemographics), average ageAtScan1 (from PNC timepoint 1) and visitagemonths (from grmpy timepoint 2) to get an age midpoint. Here, the variable has been named "midpointagemonths"

###Using these age midpoints, separate the subjects into quartiles:

```{r}
#Creating age midpoint

grmpyTemplateDemographics$midpointagemonths = (grmpyTemplateDemographics$ageAtScan1 + grmpyTemplateDemographics$visitagemonths) / 2

#Separating subjects into quartiles

grmpyTemplateDemographics$quartile <- ntile(grmpyTemplateDemographics$midpointagemonths, 4)
grmpyTemplateDemographics[order(grmpyTemplateDemographics$quartile),]

```

###You can make a histogram of these age midpoints to visualize the distribution:

```{r}
hist(grmpyTemplateDemographics$midpointagemonths)
```

For the grmpyTemplate, we want to select 32 subjects (n = 8 from each age quartile with 4 from each sex) as randomly as possible.

###Random selection extracted the following IDs:
sub-088770
sub-098567
sub-120562
sub-130121
sub-093051
sub-108202
sub-093278
sub-130896
sub-088608
sub-088773
sub-089279
sub-106880
sub-090077
sub-112332
sub-125535
sub-116354
sub-092927
sub-093734
sub-122522
sub-127008
sub-129354
sub-086350
sub-104785
sub-119302
sub-081760
sub-083999
sub-098871
sub-122528
sub-081725
sub-090262
sub-104059
sub-116019

###Now that subjects have been chosen, the nifti file for each ID needs to be organized. 

antsMultivatiateTemplateConstruction.sh requires a file containing the datapaths to your nifti files. For the grmpyTemplate, this was a text file that looked like:

/home/kmcgaughey/grmpy_midpointtemplate/sub-081725_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-081760_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-083999_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-086350_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-088608_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-088770_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-088773_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-089279_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-090077_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-090262_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-092927_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-093051_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-093278_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-093734_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-098567_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-098871_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-104059_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-104785_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-106880_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-108202_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-112332_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-116019_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-116354_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-119302_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-120562_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-122522_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-122528_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-125535_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-127008_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-129354_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-130121_T_templateBrainExtractionBrain.nii.gz
/home/kmcgaughey/grmpy_midpointtemplate/sub-130896_T_templateBrainExtractionBrain.nii.gz

**NOTE: Here, T_templateBrainExtractionBrain.nii.gz was used. While antsMultivariateTemplateConstruction.sh does have bias correction built in, future templates should still pull from the T_templateExtractedBrain0N4.nii.gz file.



###So all datapaths would be the same, nifti files were copied into a temp directory:

general=$(cat /home/kmcgaughey/grmpy_midpointtemplate/midpointtemplate_IDs.txt)

for i in $general;do
    
    bblIDs=$(echo ${i});
    echo $bblIDs
    cp /data/joy/BBL/studies/grmpy/BIDS/derivatives/antsct/${bblIDs}/${bblIDs}_T1w_SingleSubjectTemplate/T_templateBrainExtractionBrain.nii.gz /home/kmcgaughey/grmpy_midpointtemplate/${bblIDs}_T_templateBrainExtractionBrain.nii.gz 
  
done



###antsMultivatiateTemplateConstruction.sh has a number of flags that can be appended. 

-d: 3, in most cases, specifying that data is 3-dimensional
-k: Number of modalities used to construct the template (for example, if you wanted to create a multimodal template with T1, T1, and FA components "-k 3")
-c: Specifying 1 takes advantage of computing cluster. Specifying 0 will run jobs serially. 
-r: Specify 1 if there is no initial template
-s: Type of similarity metric used for registration 
-o: OutputPrefix; this will be appended to all output files (files will be named as OutputPrefix_template0.nii.gz, OutputPrefix_template1.nii.gz, … etc)

Within the output, transform files should exist for each input file, describing the transformation for that subject to the group average. The GenericAffine.mat file describes the affine transform, Warp.nii.gz describes the non linear transform, and InverseWarp.nii.gz describes the inverse of the non linear transform.



###The finalized script run for the grmpyTemplate was:

path=/home/kmcgaughey/grmpy_midpointtemplate 

/share/apps/ANTs/2014-06-23/build/bin/antsMultivariateTemplateConstruction.sh -d 3 -k 1 -c 1 -r 1 -s CC -o $path/ANTs_grmpyTemplate/grmpy_midpointTemplate $path/template_datapaths.txt



###Template-specific grmpyTemplate output was:

grmpy_midpointTemplatetemplate0Affine.txt  
grmpy_midpointTemplatetemplate0.nii.gz
grmpy_midpointTemplatetemplate0warp.nii.gz
grmpy_midpointTemplatetemplatewarplog.txt

These files have been relocated to a temp. directory: /data/joy/BBL/studies/grmpy/grmpyTemplate